BOLD Parameter Estimation Using Particle Filters

### 2010-09-02 doc/thesis/results.tex
fixed all the acronyms, think I might be done

\begin{figure}
\centering
\includegraphics[clip=true,trim=1cm 0cm 0cm 0cm, width=17cm]{images/long_converge}
\caption[\ac{BOLD} estimate converging for a very long \ac{fMRI} run.]{\ac{BOLD} estimate 
converging for a 2400 second \ac{fMRI} run. Darker bars indicate
bins with more particles with \ac{BOLD} estimate in that range. ($\sigma_x = 0.0005$, $\sigma_y = 0.001$)}
\label{fig:long_converge}
\end{figure}

\begin{figure}
\centering
\includegraphics[clip=true,trim=1cm 0cm 0cm 0cm, width=17cm]{images/long_converge_500}
\caption[First 500 seconds of the \ac{BOLD} estimate converging for a very 
long \ac{fMRI} run.]{First 500 seconds of \ac{BOLD} estimate 
converging for a 2400 second \ac{fMRI} run. Darker bars indicate
bins with more particles with \ac{BOLD} estimate in that range. ($\sigma_x = 0.0005$, $\sigma_y = 0.001$)}
\label{fig:long_converge_500}
\end{figure}

@@

\begin{table}[t]
\centering
\begin{tabular}{|c | c | c | c | c | c | c | c | c | c |}
\hline
$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  & $ \sum \tau $ & \acs{RMSR} & \acs{RMSE} \\
\hline
\rowcolor[gray]{.8}
1.45 & 0.3 & 0.47 & 0.044 & 1.94 & 1.99 & 1.8  & 5.38 &  & \\
\hline
\hline
1.2221 & 0.3449 & 0.3346 & 0.0714 & 1.6045 & 2.2753 & 1.5945 & 5.1019 &  0.003211  & 0.00224\\
1.3749 & 0.3318 & 0.3630 & 0.0733 & 1.6408 & 2.1030 & 1.5763 & 5.1187 &  0.003055  & 0.00223\\
1.1660 & 0.3221 & 0.3406 & 0.0822 & 1.6477 & 2.3535 & 1.2452 & 5.1672 &  0.003289  & 0.00205\\
1.2318 & 0.3271 & 0.3403 & 0.0796 & 1.6270 & 2.1852 & 1.3033 & 5.0439 &  0.002847  & 0.00147\\
1.1832 & 0.3179 & 0.3472 & 0.0821 & 1.5496 & 2.2912 & 1.2782 & 5.0240 &  0.003006  & 0.00213\\
1.1424 & 0.334  & 0.3473 & 0.0737 & 1.6221 & 2.2908 & 1.4025 & 5.0553 &  0.002833  & 0.00184\\
1.3004 & 0.3596 & 0.3564 & 0.0768 & 1.5641 & 2.1323 & 1.6034 & 4.9968 &  0.003028  & 0.00255\\
1.2401 & 0.3460 & 0.3398 & 0.0891 & 1.6499 & 2.2366 & 1.2900 & 5.1265 &  0.003044  & 0.00238\\
1.1709 & 0.3274 & 0.3464 & 0.0826 & 1.5373 & 2.2826 & 1.3783 & 4.9909 &  0.003345  & 0.0027 \\
1.1897 & 0.3434 & 0.3355 & 0.0798 & 1.5358 & 2.3075 & 1.4277 & 5.0330 &  0.003175  & 0.00244\\
1.184 &  0.3405 & 0.3502 & 0.0892 & 1.6103 & 2.2793 & 1.1645 & 5.0735 &  0.002889  & 0.00188\\
\hline                                                                               
1.2187 & 0.3359 & 0.3456 & 0.0800 & 1.599 & 2.2488 & 1.3876 & 5.0665 & 0.003066     & 0.00217\\
\hline
\end{tabular}
\caption{Estimated Parameters on 11 different runs with low noise. First row is the true value, 
last is the average. $\sum \tau$ is the sum of the estimated time constants.}
\label{tab:LowNoiseResults}
\end{table}

@@

\begin{table}[t]
\centering
\begin{tabular}{|c | c | c | c | c | c | c | c | c |}
\hline
$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  & \acs{RMSR} \\
\hline
1.0324 & 0.33211 & 0.34058 & 0.03012 & 1.40665 & 2.52079 & 0.5311 &   0.00167  \\
 0.98189 & 0.33047 & 0.3386 & 0.03014 & 1.45707 & 2.47232 & 0.45049 & 0.00159   \\
 1.0429 & 0.33224 & 0.34124 & 0.02946 & 1.4618 & 2.49245 & 0.43012 &  0.00165   \\
 1.02054 & 0.3321 & 0.33484 & 0.02586 & 1.45848 & 2.48741 & 0.4193 &  0.00151   \\
 1.0565 & 0.33405 & 0.33758 & 0.02791 & 1.43784 & 2.52545 & 0.47517 & 0.00152   \\
 1.01867 & 0.33528 & 0.33918 & 0.02782 & 1.48345 & 2.49605 & 0.44209 &0.00156   \\
 1.051 & 0.33038 & 0.33837 & 0.02985 & 1.47651 & 2.48621 & 0.42719 &  0.00159   \\
 1.00281 & 0.32929 & 0.33988 & 0.0298 & 1.43519 & 2.49256 & 0.48899 & 0.00164   \\
 1.00893 & 0.33273 & 0.33982 & 0.0289 & 1.42903 & 2.49754 & 0.45688 & 0.00168   \\
 1.01289 & 0.33275 & 0.3376 & 0.02997 & 1.41188 & 2.49881 & 0.50628 & 0.00183   \\
 1.10247 & 0.33371 & 0.3419 & 0.02939 & 1.43774 & 2.53384 & 0.44079 & 0.00195   \\
\hline
1.03009 & 0.33228 & 0.33905 & 0.02902 & 1.44506 & 2.50031 & 0.46076 & 0.00165 \\
\hline
\end{tabular}
\caption{Estimated Parameters on 11 different runs with low noise and no signal present.
 ($\sigma_x = 0.005$, $\sigma_y = 0.01$)}
\label{tab:NoiseOnlyResults}
\end{table}

@@

\subsection{Single Voxel Summary}
\label{sec:SingleVoxelReview}
Because of the variability in the signal levels, the
raw \ac{RMSR} cannot be used to rate the fit. As demonstrated by \autoref{tab:NoiseOnlyResults},
a low \ac{RMSR} does not necessarily indicate a good fit.
Therefore, a normalized version of the \ac{RMSR} was used. To normalize the
residual, both the estimated and preprocessed \ac{BOLD} signals were divided by an estimator
of scale and then the new \ac{RMSR} was calculated.
Considering the tendency of \ac{fMRI} to have large unexplainable
peaks and troughs, the \ac{MAD} was used (\autoref{eq:mad}).
This is an estimator of the standard
deviation, and thus a good estimator of the scale of the input signal.
The normalized \ac{RMSR} values are
shown in \autoref{tab:SingleVoxelActivationComparison}. A second potential method of
gauging performance is mutual information. Mutual information is a method of measuring
the interdependence of two random variables. If two signals are truly independent,
then the mutual information will be zero. Although ideally suited to discrete
distributions, by using histograms
it is possible to derive a joint distribution of two signals. The algorithm for
mutual information is based on that joint distribution:
\begin{equation}
\sum_{x,y} p(x,y) \log_2\left(\frac{p(x,y)}{p(x)p(y)}\right)
\end{equation}
Unfortunately the number of bins causes bias in the output, thus to correct for this,
I subtracted the estimated bias:
\begin{equation}
\text{bias} = \frac{N_{bins}}{2Nlog(2)}
\end{equation}
where $N$ is the number of samples and $N_{bins}$ is the number of bins. For all
the mutual information estimates in this work 6 bins were used for the marginal
distribution of each signal. Additionally, throughout log base 2 will be used.
This leads to 36 total bins in the joint, so the bias is:
\begin{equation}
\text{bias} = \frac{18}{N}
\end{equation}
 Note that subtracting the bias can result in negative mutual
information, which should not technically be possible; so any negative mutual information
was taken as 0.

\begin{table}[t]
\centering
\begin{tabular}{|c | c | c | c | c | c | c | c | c |}
\hline
& \multicolumn{4}{|c|}{Signal} & \multicolumn{4}{|c|}{No Signal}\\
\hline
& \multicolumn{2}{|c|}{Low Noise} & \multicolumn{2}{|c|}{High Noise}
& \multicolumn{2}{|c|}{$\sigma_y = 0.001, \sigma_x = 0.0005$}
& \multicolumn{2}{|c|}{$\sigma_y = 0.01, \sigma_x = 0.005$}\\
\hline
& $M.I.$ & N. Res. &
  $M.I.$ & N. Res. &
  $M.I.$ & N. Res. &
  $M.I.$ & N. Res. \\
\hline
\hline
1 &   0.86687  &0.47801 &  0.09077  &1.03894 &  0.06326  &  1.29501 & 0.03024  &1.33641 \\
2 &   0.93975  &0.53177 &  0.13767  &0.95165 &  -0.01075  & 1.30175 & -0.02677 &1.33667 \\
3 &   0.82382  &0.5458  &  0.13505  &0.99539 &  0.02345  &  1.26287 & -0.0111  &1.15957 \\
4 &   0.94661  &0.49824 &  0.04341  &1.16129 &  -0.00906  & 1.43196 & 0.00147  &1.09988 \\
5 &   0.94281  &0.46805 &  0.13718  &1.03972 &  0.00663  &  1.25664 & -0.00204 &1.20107 \\
6 &   0.92539  &0.459   &  0.12337  &1.00214 &  -0.00816  & 1.2708 &  0.01775  &1.04589 \\
7 &   0.98892  &0.46096 &  0.15381  &1.08847 &  0.02664  &  1.15441 & 0.03163  &1.20543 \\
8 &   0.98796  &0.51838 &  0.11325  &1.05962 &  0.03285  &  1.27456 & 0.01951  &1.1225 \\
9 &   0.8804   &0.5253  &  0.09669  &1.0157  &  0.01628  &  1.32024 & 0.01039  &1.08637 \\
10 &  0.88721   &0.49211 & 0.18339  &1.18996 &  0.00407  &  1.34456 & 0.00508  &1.22135 \\
11 &  0.96644   &0.49092 & 0.10949  &0.95368 &  0.03323  &  1.32522 & -0.01284 &1.11737 \\
\hline
mean &0.92329  &0.49714 & 0.12037   &1.04514 &  0.01622  &  1.29437 & 0.00576  &1.17568 \\
\hline
min &  0.82382  &0.459   &  0.04341  &0.95165 & -0.01075  & 1.15441 & -0.02677 &1.04589 \\
\hline
max &  0.98892  &0.5458  & 0.18339   &1.18996 & 0.06326  &  1.43196 & 0.03163  &1.33667 \\
\hline
\end{tabular}
\caption{Mutual Information and the normalized \ac{RMSE}, for each of the previous
sections.}
\label{tab:SingleVoxelActivationComparison}
\end{table}

Comparing the results of \autoref{sec:SimHighNoise} and 
\autoref{sec:PureNoiseHighMag}
in \autoref{tab:SingleVoxelActivationComparison},
distinguishing between these cases with either normalized \ac{RMSR} or 
mutual information is not clear cut. While the average mutual information 
is more than 10 times
the average mutual information in the two non-signal cases, the maximum
mutual information of the low noise/no signal case exceeds the minimum M.I. 
of the high noise/signal
case. The Low Noise/signal determination is easier to make; given
the minimum mutual information is above $.8$ and the maximum normalized 
\ac{RMSR} is below $.6$.
However, it is worth noting that the worst case scenario  for mutual information (maximum) in the
low noise/no signal case does not coincide with the worse case (minimum) normalized residual.
There is no reason why this has to be the case, but it could be beneficial. In
other words, if it were necessary to make a statement that
a particular voxel were active or inactive; the accuracy would improve if both techniques
were used with loose restrictions. In all, both M.I. and the normalized residual
provide a good measure of performance.

There were two primary purposes of these tests. First, given the nature of Monte Carlo
techniques it is important to ensure consistency of results. Although the parameter
sets were inconsistent, the quality of the fit for the \ac{BOLD} output was often
accurate even when noise was drastically increased. The second purpose was to validate
Mutual Information and the normalized \ac{RMSR} as measurements of output quality.
Although not perfect, both methods do provide a decent metric.

@@

\subfigure[Normalized \ac{RMSR} Map. Scale is Normalized \ac{RMSE}. Lower (yellow) is better.]
{\label{fig:simslice_hm_res} \includegraphics[height=8.5cm]{images/sim_hm}}
\subfigure[Mutual Information Map. Scale is bits. Higher (yellow) indicates better fit.]
{\label{fig:sim_hm_mi} \includegraphics[height=8.5cm]{images/sim_hm_mi}}

\caption{Comparison of activation regions, a) grey matter and regions, b) \ac{SNR} of active
regions, c) \acl{RMSR} Map, d) Mutual Information Map.}
\label{fig:simslice_hm}
\end{figure}

@@

Note that region 4 had a very low $\epsilon$, putting it below the noise
threshold. For this reason, the only areas
with significant estimates of the \ac{BOLD} time series were 1,2 and 3.
Notice that the
regions 1, 2 an 3 stick out in both the \ac{RMSR} and the mutual information
map, indicating that the particle filter was successful in matching those regions.
Mutual information was an extremely successful metric, with the exception of
a few false positives (\autoref{fig:simslice_hm_res}). 
\autoref{fig:sim_hm2} shows what the heat map looks
like when the threshold is raised from $0.1$ to $0.15$.
\begin{figure}
\centering
\subfigure{
\includegraphics[width=10cm]{images/sim_hm_15_6.png}}
\subfigure{\includegraphics[width=1cm]{images/scale6.png}}
\caption{More stringent mutual information heatmap. Higher (yellow) is better.}
\label{fig:sim_hm2}
\end{figure}

\begin{figure} %bottom left
\centering
\includegraphics[clip=true,trim=2.5cm 2cm 2cm 1cm,width=\textwidth]{images/slicesim_hist1}
\caption[Histogram of estimated parameters in section 1]{Histogram of estimated parameters in 
section 1 in voxels 
with mutual information greater than $0.15$} \label{fig:slicesim_hist1}
\end{figure}

\begin{figure} %top left
\centering
\includegraphics[clip=true,trim=2.5cm 2cm 2cm 1cm,width=\textwidth]{images/slicesim_hist2}
\caption[Histogram of estimated parameters in section 2]
{Histogram of estimated parameters in section 2 in voxels with mutual information greater
than $0.15$}
\label{fig:slicesim_hist2}
\end{figure}

\begin{figure} %top right
\centering
\includegraphics[clip=true,trim=2.5cm 2cm 2cm 1cm,width=\textwidth]{images/slicesim_hist3}
\caption[Histogram of estimated parameters in section 3]
{Histogram of estimated parameters in section 3 in voxels with mutual information greater
than $0.15$}
\label{fig:slicesim_hist3}
\end{figure}

If the thresholds are left at the values of \autoref{fig:simslice_hm}, then
the normalized \ac{RMSR} gave $2/1479$  false positives and mutual information gave
$11/1479$. However, normalized \ac{RMSR} also clearly missed some active voxels, although
false negatives are harder to quantify because of the presence of white matter.
Raising the threshold to $0.15$ eliminated the false positives in the
Mutual Information map, although some active voxels were also removed.

The histograms again demonstrate that a single point estimate of the parameters
is elusive for this set of parameters. However the data clearly show the power
of the particle filter at identifying regions of activation.
The thresholds applied to this slice, both for mutual information and
\acp{RMSR} are arbitrary. Applying a
threshold is helpful for visualization but is rarely useful for further analysis.
As noted in \autoref{sec:SingleVoxelReview},
the false positives present in the Mutual Information
map are different from those in the \ac{RMSR} map. This furthers the argument
for combining the two metrics to increase power. Although
at first glance it would appear that there are false negatives in
\autoref{fig:sim_hm_mi}; this is
not actually the case. \ac{POSSUM} simulates different tissues, and white matter
does not typically  have a \ac{BOLD} response. This explains some of the 
holes in regions
2 and 3. These results certainly indicate that the particle filter is effective
at regressing against a noisy signal.

>>>>

\begin{figure}
\centering
\includegraphics[clip=true,trim=1cm 0cm 0cm 0cm, width=17cm]{images/long_converge}
\caption[\acs{BOLD} estimate converging for a very long \acs{fMRI} run.]{\acs{BOLD} estimate 
converging for a 2400 second \acs{fMRI} run. Darker bars indicate
bins with more particles with \acs{BOLD} estimate in that range. ($\sigma_x = 0.0005$, $\sigma_y = 0.001$)}
\label{fig:long_converge}
\end{figure}

\begin{figure}
\centering
\includegraphics[clip=true,trim=1cm 0cm 0cm 0cm, width=17cm]{images/long_converge_500}
\caption[First 500 seconds of the \acs{BOLD} estimate converging for a very 
long \acs{fMRI} run.]{First 500 seconds of \acs{BOLD} estimate 
converging for a 2400 second \acs{fMRI} run. Darker bars indicate
bins with more particles with \acs{BOLD} estimate in that range. ($\sigma_x = 0.0005$, $\sigma_y = 0.001$)}
\label{fig:long_converge_500}
\end{figure}

@@

\begin{table}[t]
\centering
\begin{tabular}{|c | c | c | c | c | c | c | c | c | c |}
\hline
$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  & $ \sum \tau $ & \acs{RMSR} & \acs{RMSE} \\
\hline
\rowcolor[gray]{.8}
1.45 & 0.3 & 0.47 & 0.044 & 1.94 & 1.99 & 1.8  & 5.38 &  & \\
\hline
\hline
1.2221 & 0.3449 & 0.3346 & 0.0714 & 1.6045 & 2.2753 & 1.5945 & 5.1019 &  0.003211  & 0.00224\\
1.3749 & 0.3318 & 0.3630 & 0.0733 & 1.6408 & 2.1030 & 1.5763 & 5.1187 &  0.003055  & 0.00223\\
1.1660 & 0.3221 & 0.3406 & 0.0822 & 1.6477 & 2.3535 & 1.2452 & 5.1672 &  0.003289  & 0.00205\\
1.2318 & 0.3271 & 0.3403 & 0.0796 & 1.6270 & 2.1852 & 1.3033 & 5.0439 &  0.002847  & 0.00147\\
1.1832 & 0.3179 & 0.3472 & 0.0821 & 1.5496 & 2.2912 & 1.2782 & 5.0240 &  0.003006  & 0.00213\\
1.1424 & 0.334  & 0.3473 & 0.0737 & 1.6221 & 2.2908 & 1.4025 & 5.0553 &  0.002833  & 0.00184\\
1.3004 & 0.3596 & 0.3564 & 0.0768 & 1.5641 & 2.1323 & 1.6034 & 4.9968 &  0.003028  & 0.00255\\
1.2401 & 0.3460 & 0.3398 & 0.0891 & 1.6499 & 2.2366 & 1.2900 & 5.1265 &  0.003044  & 0.00238\\
1.1709 & 0.3274 & 0.3464 & 0.0826 & 1.5373 & 2.2826 & 1.3783 & 4.9909 &  0.003345  & 0.0027 \\
1.1897 & 0.3434 & 0.3355 & 0.0798 & 1.5358 & 2.3075 & 1.4277 & 5.0330 &  0.003175  & 0.00244\\
1.184 &  0.3405 & 0.3502 & 0.0892 & 1.6103 & 2.2793 & 1.1645 & 5.0735 &  0.002889  & 0.00188\\
\hline                                                                               
1.2187 & 0.3359 & 0.3456 & 0.0800 & 1.599 & 2.2488 & 1.3876 & 5.0665 & 0.003066     & 0.00217\\
\hline
\end{tabular}
\caption[Estimated Parameters with low noise.]
{Estimated Parameters on 11 different runs with low noise. First row is the true value, 
last is the average. $\sum \tau$ is the sum of the estimated time constants.}
\label{tab:LowNoiseResults}
\end{table}

@@

\begin{table}[t]
\centering
\begin{tabular}{|c | c | c | c | c | c | c | c | c |}
\hline
$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  & \acs{RMSR} \\
\hline
1.0324 & 0.33211 & 0.34058 & 0.03012 & 1.40665 & 2.52079 & 0.5311 &   0.00167  \\
 0.98189 & 0.33047 & 0.3386 & 0.03014 & 1.45707 & 2.47232 & 0.45049 & 0.00159   \\
 1.0429 & 0.33224 & 0.34124 & 0.02946 & 1.4618 & 2.49245 & 0.43012 &  0.00165   \\
 1.02054 & 0.3321 & 0.33484 & 0.02586 & 1.45848 & 2.48741 & 0.4193 &  0.00151   \\
 1.0565 & 0.33405 & 0.33758 & 0.02791 & 1.43784 & 2.52545 & 0.47517 & 0.00152   \\
 1.01867 & 0.33528 & 0.33918 & 0.02782 & 1.48345 & 2.49605 & 0.44209 &0.00156   \\
 1.051 & 0.33038 & 0.33837 & 0.02985 & 1.47651 & 2.48621 & 0.42719 &  0.00159   \\
 1.00281 & 0.32929 & 0.33988 & 0.0298 & 1.43519 & 2.49256 & 0.48899 & 0.00164   \\
 1.00893 & 0.33273 & 0.33982 & 0.0289 & 1.42903 & 2.49754 & 0.45688 & 0.00168   \\
 1.01289 & 0.33275 & 0.3376 & 0.02997 & 1.41188 & 2.49881 & 0.50628 & 0.00183   \\
 1.10247 & 0.33371 & 0.3419 & 0.02939 & 1.43774 & 2.53384 & 0.44079 & 0.00195   \\
\hline
1.03009 & 0.33228 & 0.33905 & 0.02902 & 1.44506 & 2.50031 & 0.46076 & 0.00165 \\
\hline
\end{tabular}
\caption[Estimated Parameters with low noies, and no signal.]
{Estimated Parameters on 11 different runs with low noise and no signal present.
 ($\sigma_x = 0.005$, $\sigma_y = 0.01$)}
\label{tab:NoiseOnlyResults}
\end{table}

@@

\subsection{Single Voxel Summary}
\label{sec:SingleVoxelReview}
Because of the variability in the signal levels, the
raw \ac{RMSR} cannot be used to rate the fit. As demonstrated by \autoref{tab:NoiseOnlyResults},
a low \ac{RMSR} does not necessarily indicate a good fit.
Therefore, a normalized version of the \ac{RMSR} was used. To normalize the
residual, both the estimated and preprocessed \ac{BOLD} signals were divided by an estimator
of scale and then the new \ac{RMSR} was calculated.
Considering the tendency of \ac{fMRI} to have large unexplainable
peaks and troughs, the \ac{MAD} was used (\autoref{eq:mad}).
This is an estimator of the standard
deviation, and thus a good estimator of the scale of the input signal.
The normalized \ac{RMSR} values are
shown in \autoref{tab:SingleVoxelActivationComparison}. A second potential method of
gauging performance is \ac{MI}. \acl{MI} is a method of measuring
the interdependence of two random variables. If two signals are truly independent,
then the \ac{MI} will be zero. Although ideally suited to discrete
distributions, by using histograms
it is possible to derive a joint distribution of two signals. The algorithm for
\ac{MI} is based on that joint distribution:
\begin{equation}
\sum_{x,y} p(x,y) \log_2\left(\frac{p(x,y)}{p(x)p(y)}\right)
\end{equation}
Unfortunately the number of bins causes bias in the output, thus to correct for this,
I subtracted the estimated bias:
\begin{equation}
\text{bias} = \frac{N_{bins}}{2Nlog(2)}
\end{equation}
where $N$ is the number of samples and $N_{bins}$ is the number of bins. For all
the \ac{MI} estimates in this work 6 bins were used for the marginal
distribution of each signal. Additionally, throughout log base 2 will be used.
This leads to 36 total bins in the joint, so the bias is:
\begin{equation}
\text{bias} = \frac{18}{N}
\end{equation}
 Note that subtracting the bias can result in negative 
\ac{MI}, which should not technically be possible; so any negative \acl{MI}
was taken as 0.

\begin{table}[t]
\centering
\begin{tabular}{|c | c | c | c | c | c | c | c | c |}
\hline
& \multicolumn{4}{|c|}{Signal} & \multicolumn{4}{|c|}{No Signal}\\
\hline
& \multicolumn{2}{|c|}{Low Noise} & \multicolumn{2}{|c|}{High Noise}
& \multicolumn{2}{|c|}{$\sigma_y = 0.001, \sigma_x = 0.0005$}
& \multicolumn{2}{|c|}{$\sigma_y = 0.01, \sigma_x = 0.005$}\\
\hline
& $MI$ & N. Res. &
  $MI$ & N. Res. &
  $MI$ & N. Res. &
  $MI$ & N. Res. \\
\hline
\hline
1 &   0.86687  &0.47801 &  0.09077  &1.03894 &  0.06326  &  1.29501 & 0.03024  &1.33641 \\
2 &   0.93975  &0.53177 &  0.13767  &0.95165 &  -0.01075  & 1.30175 & -0.02677 &1.33667 \\
3 &   0.82382  &0.5458  &  0.13505  &0.99539 &  0.02345  &  1.26287 & -0.0111  &1.15957 \\
4 &   0.94661  &0.49824 &  0.04341  &1.16129 &  -0.00906  & 1.43196 & 0.00147  &1.09988 \\
5 &   0.94281  &0.46805 &  0.13718  &1.03972 &  0.00663  &  1.25664 & -0.00204 &1.20107 \\
6 &   0.92539  &0.459   &  0.12337  &1.00214 &  -0.00816  & 1.2708 &  0.01775  &1.04589 \\
7 &   0.98892  &0.46096 &  0.15381  &1.08847 &  0.02664  &  1.15441 & 0.03163  &1.20543 \\
8 &   0.98796  &0.51838 &  0.11325  &1.05962 &  0.03285  &  1.27456 & 0.01951  &1.1225 \\
9 &   0.8804   &0.5253  &  0.09669  &1.0157  &  0.01628  &  1.32024 & 0.01039  &1.08637 \\
10 &  0.88721   &0.49211 & 0.18339  &1.18996 &  0.00407  &  1.34456 & 0.00508  &1.22135 \\
11 &  0.96644   &0.49092 & 0.10949  &0.95368 &  0.03323  &  1.32522 & -0.01284 &1.11737 \\
\hline
mean &0.92329  &0.49714 & 0.12037   &1.04514 &  0.01622  &  1.29437 & 0.00576  &1.17568 \\
\hline
min &  0.82382  &0.459   &  0.04341  &0.95165 & -0.01075  & 1.15441 & -0.02677 &1.04589 \\
\hline
max &  0.98892  &0.5458  & 0.18339   &1.18996 & 0.06326  &  1.43196 & 0.03163  &1.33667 \\
\hline
\end{tabular}
\caption{\acs{MI} and the normalized \acs{RMSE}, for each of the previous sections.}
\label{tab:SingleVoxelActivationComparison}
\end{table}

Comparing the results of \autoref{sec:SimHighNoise} and 
\autoref{sec:PureNoiseHighMag}
in \autoref{tab:SingleVoxelActivationComparison},
distinguishing between these cases with either normalized \ac{RMSR} or 
\ac{MI} is not clear cut. While the average \ac{MI}
is more than 10 times
the average \ac{MI} in the two non-signal cases, the maximum
\ac{MI} of the low noise/no signal case exceeds the minimum \ac{MI}
of the high noise/signal
case. The Low Noise/signal determination is easier to make; given
the minimum \ac{MI} is above $.8$ and the maximum normalized 
\ac{RMSR} is below $.6$.
However, it is worth noting that the worst case scenario  for \ac{MI} (maximum) in the
low noise/no signal case does not coincide with the worse case (minimum) normalized residual.
There is no reason why this has to be the case, but it could be beneficial. In
other words, if it were necessary to make a statement that
a particular voxel were active or inactive; the accuracy would improve if both techniques
were used with loose restrictions. In all, both \ac{MI} and the normalized residual
provide a good measure of performance.

There were two primary purposes of these tests. First, given the nature of Monte Carlo
techniques it is important to ensure consistency of results. Although the parameter
sets were inconsistent, the quality of the fit for the \ac{BOLD} output was often
accurate even when noise was drastically increased. The second purpose was to validate
\ac{MI} and the normalized \ac{RMSR} as measurements of output quality.
Although not perfect, both methods do provide a decent metric.

@@

\subfigure[Normalized \ac{RMSR} Map. Scale is Normalized \ac{RMSE}. Lower (yellow) is better.]
{\label{fig:simslice_hm_res} \includegraphics[height=8.5cm]{images/sim_hm}}
\subfigure[\ac{MI} Map. Scale is bits. Higher (yellow) indicates better fit.]
{\label{fig:sim_hm_mi} \includegraphics[height=8.5cm]{images/sim_hm_mi}}

\caption[Comparison of activation regions.]
{Comparison of activation regions, a) grey matter and regions, b) \acs{SNR} of active
regions, c) \acs{RMSR} Map, d) \acs{MI} Map.}
\label{fig:simslice_hm}
\end{figure}

@@

Note that region 4 had a very low $\epsilon$, putting it below the noise
threshold. For this reason, the only areas
with significant estimates of the \ac{BOLD} time series were 1,2 and 3.
Notice that the
regions 1, 2 an 3 stick out in both the \ac{RMSR} and the \ac{MI}
map, indicating that the particle filter was successful in matching those regions.
\ac{MI} was an extremely successful metric, with the exception of
a few false positives (\autoref{fig:simslice_hm_res}). 
\autoref{fig:sim_hm2} shows what the heat map looks
like when the threshold is raised from $0.1$ to $0.15$.
\begin{figure}
\centering
\subfigure{
\includegraphics[width=10cm]{images/sim_hm_15_6.png}}
\subfigure{\includegraphics[width=1cm]{images/scale6.png}}
\caption{More stringent \acs{MI} heatmap. Higher (yellow) is better.}
\label{fig:sim_hm2}
\end{figure}

\begin{figure} %bottom left
\centering
\includegraphics[clip=true,trim=2.5cm 2cm 2cm 1cm,width=\textwidth]{images/slicesim_hist1}
\caption[Histogram of estimated parameters in section 1]{Histogram of estimated parameters in 
section 1 in voxels 
with \acs{MI} greater than $0.15$} \label{fig:slicesim_hist1}
\end{figure}

\begin{figure} %top left
\centering
\includegraphics[clip=true,trim=2.5cm 2cm 2cm 1cm,width=\textwidth]{images/slicesim_hist2}
\caption[Histogram of estimated parameters in section 2]
{Histogram of estimated parameters in section 2 in voxels with \acs{MI} greater
than $0.15$}
\label{fig:slicesim_hist2}
\end{figure}

\begin{figure} %top right
\centering
\includegraphics[clip=true,trim=2.5cm 2cm 2cm 1cm,width=\textwidth]{images/slicesim_hist3}
\caption[Histogram of estimated parameters in section 3]
{Histogram of estimated parameters in section 3 in voxels with \acs{MI} greater
than $0.15$}
\label{fig:slicesim_hist3}
\end{figure}

If the thresholds are left at the values of \autoref{fig:simslice_hm}, then
the normalized \ac{RMSR} gave $2/1479$  false positives and \ac{MI} gave
$11/1479$. However, normalized \ac{RMSR} also clearly missed some active voxels, although
false negatives are harder to quantify because of the presence of white matter.
Raising the threshold to $0.15$ eliminated the false positives in the
\ac{MI} map, although some active voxels were also removed.

The histograms again demonstrate that a single point estimate of the parameters
is elusive for this set of parameters. However the data clearly show the power
of the particle filter at identifying regions of activation.
The thresholds applied to this slice, both for \ac{MI} and
\acp{RMSR} are arbitrary. Applying a
threshold is helpful for visualization but is rarely useful for further analysis.
As noted in \autoref{sec:SingleVoxelReview},
the false positives present in the \ac{MI}
map are different from those in the \ac{RMSR} map. This furthers the argument
for combining the two metrics to increase power. Although
at first glance it would appear that there are false negatives in
\autoref{fig:sim_hm_mi}; this is
not actually the case. \ac{POSSUM} simulates different tissues, and white matter
does not typically  have a \ac{BOLD} response. This explains some of the 
holes in regions
2 and 3. These results certainly indicate that the particle filter is effective
at regressing against a noisy signal.
